---
title: "Homework 1"
---

```{r, echo=FALSE, include=FALSE}
library(kableExtra)
library(knitr)
library(tidyverse)
```



### 1. OLS Estimation

  a. Using the scalar formulation of the ANOVA model $y_{ij} \sim N\left(\mu_j,\sigma^2 \right)$, show that $\widehat{\mu}_{OLS}=(\overline{y}_1,\cdots,\overline{y}_J)$, where $\overline{y}_j$ is the sample mean in group $j$.
  
  b. Write the model in (a) in matrix form assuming you have 3 groups with 3 observations each and show by hand (or typed!) that $(X'X)^{-1}X'y$ yields the same estimates.
  
  c. Using the scalar formulation of the ANOVA model $y_{ij} \sim N\left(\mu+\alpha_j,\sigma^2 \right)$ with the constraint $\sum_j \alpha_j=0$, show
    
    i. $\widehat{\mu}=\overline{y}_{\cdot \cdot}$, where $\overline{y}_{\cdot \cdot}$ is the grand mean over all observations
    
    ii. $\widehat{\mu}=\frac{1}{J}\sum_j \widehat{\mu}_j$ when the sample sizes in each group j $n_j=n$ for all groups
    
    iii. $\widehat{\alpha}_j=\widehat{\mu}_j-\widehat{\mu}=\overline{y}_{j}-\overline{y}_{\cdot \cdot}$
    
    
### 2. Contrasts

Suppose you fit an ANOVA model to responses collected from J=3 groups, each of sample size $n$. You consider the following two ANOVA models.

\begin{eqnarray}
y_{ij}&=&\mu_j + \varepsilon_{ij} \\
y_{ij}&=& \mu + \alpha_1I(j=1)+\alpha_2I(j=2)+\varepsilon_{ij}
\end{eqnarray}

  a. Find the linear combinations of parameters in model (2) that are equivalent to $\mu_1-\mu_2$, $\mu_1-\mu_3$, and $\mu_2-\mu_3$ in model (1).
  
  b. Show that the estimates of these mean differences are identical regardless of the coding scheme used.
  
  
### 3. Game of Thrones

The handling of female characters in the American series Game of Thrones, and indeed whether it is feminist or mysogynistic, has been hotly debated. The dataset gotscreen.Rdata contains information on the number of seconds of screentime for members of each gender in each episode of 7 seasons of Game of Thrones. Using a two-way ANOVA model with interaction, explore whether screentime differs by gender (male, female, or unspecified) and whether there are any differences in potential gender effects across the 7 seasons of the show. Your answer should include the following details.

  1. A clear specification of the model using an equation, including clear specification of any modeling assumptions.
  2. Point and interval estimates for each parameter in the linear predictor of your model.
  3. Clear specification of any hypothesis tests or other inferential techniques used to evaluate the questions posed above.
  4. Evidence of adequacy of model fit and evaluation of suitability of any assumptions.
  5. Clear description of results in language accessible to the average fan of the show, including graphical displays as appropriate.
  
  
### 4. Prove or disprove. Suppose $X$ and $Y$ each follow *univariate* normal distributions and that $\text{Cov}(X,Y)=0$. Then $X$ is independent of $Y$.



``` {r gotimport}
library(readxl)
gameofthronesseason1234567 <- read_excel("~/Documents/TEACHING/STA410_610/DATA/gameofthronesseason1234567.xlsx")
View(gameofthronesseason1234567)

library(dplyr)
gotscreen=gameofthronesseason1234567 %>%
    group_by(Episode, Gender, Season) %>%
    summarize(seccount = sum(screensec))

gameofthronesseason1234567 %>%
    group_by(Gender, Season) %>%
    summarize(secmean = mean(screensec))


aov.got=aov(seccount~factor(Season)*factor(Gender),data=gotscreen)
save(gotscreen, file = "~/Documents/TEACHING/STA410_610/DATA/gotscreen.RData")

```
    
    HAVE THEM DO REFERENCE CELL CODING AND ALSO ESTIMATE CONTRASTS ON THIS HOMEWORK TOO GOING BACK AND FORTH BETWEEN CODING SCHEMES


