---
title: "Homework 2"
---

```{r, echo=FALSE, include=FALSE}
library(kableExtra)
library(knitr)
library(tidyverse)
```



### 1. Hierarchical Normal Model

a. Derive the full conditional posterior distribution of $\mu_j$ using the specification provided in the notes, showing all intermediate steps to verify its conditional mean is X and cond var is Y.

b. Derive the full conditional posterior distribution of $\sigma^2$ blah blah.



``` {r gotimport}
library(readxl)
gameofthronesseason1234567 <- read_excel("~/Documents/TEACHING/STA410_610/DATA/gameofthronesseason1234567.xlsx")
View(gameofthronesseason1234567)

library(dplyr)
gotscreen=gameofthronesseason1234567 %>%
    group_by(Episode, Gender, Season) %>%
    summarize(seccount = sum(screensec))

gameofthronesseason1234567 %>%
    group_by(Gender, Season) %>%
    summarize(secmean = mean(screensec))


aov.got=aov(seccount~factor(Season)*factor(Gender),data=gotscreen)
save(gotscreen, file = "~/Documents/TEACHING/STA410_610/DATA/gotscreen.RData")

```
    
    HAVE THEM DO REFERENCE CELL CODING AND ALSO ESTIMATE CONTRASTS ON THIS HOMEWORK TOO GOING BACK AND FORTH BETWEEN CODING SCHEMES


