---
title: Random Effects ANOVA
output: 
  revealjs::revealjs_presentation:
    theme: night
    highlight: espresso
    center: true
    transition: none
    css: styles.css
    fig_caption: true
    reveal_options:
      progress: true
      slideNumber: true
---

## Readings


This lecture is based on Section 2.2 and some of Chapter 3 of Hoff.

## Random Effects ANOVA

Random effects ANOVA is a simple hierarchical model. In this framework we assume that group-specific means are distributed around some overall mean.


We will introduce this model in the context of a study in which the groups are the 85 counties in the state of Minnesota.

We consider this study rather than just the bike data because estimates of county-specific estimates are of interest, and some counties have very few observed data points. (In the bike data, we have a substantial number of replicates for each curb distance.)

## Motivating Example: Radon Study

Radon, a naturally-occurring radioactive gas, is a carcinogen known to cause lung cancer in high concentrations.  Several thousand lung cancer deaths in the US per year are attributable to radon exposure.

Radon levels in US homes vary greatly, and some homes have dangerously high radon levels.  In order to identify areas of the US with high radon exposures, the US EPA conducted a study of radon levels in a random sample of more than 80,000 homes.

[Click here](http://www.ncradon.org/ncradon/) to check highest recorded radon levels in your area. Note that these levels may come from short-term home test kits, which vary widely in accuracy.

(This example is taken from the excellent book by Gelman and Hill.)

## Radon study

We wish to estimate the distribution of radon levels in houses $i$ within 85 counties $j$ in Minnesota. The outcome $y_i$ is the natural log of measured radon levels. 

- One estimate would be the average of all radon levels in Minnesota (same estimate for all counties), $\overline{y}_{\cdot \cdot}$, but this ignores variation across counties, and some counties may have higher radon levels naturally than others (radon is more commonly found in soils with granite rock, as opposed to some other soil types).

- Another estimate would be just to average the radon level in each
  county, $\overline{y}_j$, which can over-fit the data within county (for example, Lac Qui Parle County, which has the highest observed radon level of all the 85 MN counties, had radon measures from only 2 homes). This is similar to using an ANOVA model with a *fixed effect* for county.

## 
![Estimates and SE's for log radon levels in MN counties versus the
    (jittered) sample size. The horizontal line indicates the overall state mean (figure from Gelman and Hill).](figures/gelmannopool.jpg){width=50%}

Note we get pretty good (low variance) estimates in counties where more samples were taken, while our estimates are not great in counties where just a few samples were obtained. 

##

The figure contrasts two extreme approaches to obtaining estimates of group means $\mu_j$ in this type of setting.  A common procedure might be the following.  Fit the ANOVA model $y_{ij}=\mu+\alpha_j+\varepsilon_{ij}$, where $\varepsilon_{ij} \sim N(0,\sigma^2)$, testing the significance of the groups using an overall F test (here, an 84 degree of freedom test, as we're testing all 85 means are the same).

  - If $p<0.05$, use the estimate $\widehat{\mu}_j=\overline{y}_j$ for the mean in each county
  - If $p>0.05$, use the estimate $\widehat{\mu}_j=\overline{y}$ for the mean in each county
  
With either case, we will be using sub-optimal estimates for some counties, and the above method is fairly extreme (all or nothing one way or the other!).

## 

An improvement might be using the estimate $\overline{y}_j$ for counties with sufficient sample size and the estimate $\overline{y}$ for counties where the variability is too high.  

Important question: how do we define "sufficient" and "too high"?

## Random Effects ANOVA

*Random effects* ANOVA is a special case of a *hierarchical* or *multilevel* linear model that provides a nice framework for borrowing information across groups when needed to stabilize estimates.  We can specify such a model as $y_{ij}=\mu+\alpha_j+\varepsilon_{ij}$, where $\varepsilon_{ij} \overset{iid}{\sim} N(0,\sigma^2)$ and $\alpha_j \overset{iid}{\sim} N(0,\tau^2)$. The model on $\alpha_j$ allows us to borrow information in order to obtain better group-specific estimates when needed; because $\alpha_j$ is now viewed as random, the model can be called a *random effects* model. 

This particular model is sometimes called a *random intercept* model because each group has its own intercept, $\mu_j=\mu+\alpha_j$, that follows a Gaussian distribution.

## Random Effects ANOVA for Radon Data


![Estimates and SE's for log radon levels in MN versus the
    (jittered) sample size. The horizontal line indicates the complete pooling estimate (Gelman and Hill). The circled data point had the highest estimated mean radon level in fixed effects ANOVA.](figures/gelman1.jpg){width=50%}
    
## Radon study

The multilevel estimates in the previous slide represent a compromise between the two extremes.  In this simple setting (with no predictors), the multilevel estimate for county $j$ can be approximated as a weighted average of the mean of all observations in the county, weighting both the unpooled estimate $\overline{y}_j$ and the mean over all counties $\overline{y}$.

## 

How does random effects ANOVA borrow information?

The multilevel estimate

$$\widehat{\alpha}_j \approx
\frac{\frac{n_j}{\sigma^2}\overline{y}_j+\frac{1}{\tau^2}\overline{y}}{\frac{n_j}{\sigma^2}+\frac{1}{\tau^2}},$$
   
where 
    
- $n_j$ is the number of homes measured in county $j$

- $\sigma^2$ is the within-county variance in the log radon
  measurements
  
- $\tau^2$ is the variance across the average log radon
  levels of different counties

## 

The weighted average reflects the relative amount of information available on each individual county, compared to that available across all counties.

- Averages from counties with smaller sample sizes are less precise, so the weighting shrinks the multilevel estimates closer to the overall state average.  For example, if $n_j=0,$ the multilevel estimate is just $\overline{y}$.

- Averages from counties with larger sample sizes are more precise, and the multilevel estimates are closer to the county averages.  As $n_j \rightarrow \infty$, the multilevel estimate is just the county average $\overline{y}_j$.

- In intermediate cases, the multilevel estimate is in between the
  extremes.
  
In practice, we carry out all estimation together (estimate variances along with the mean parameters), but we won't worry about this yet.

## Understanding the Model

These estimates are often called *shrinkage estimates*, as they "shrink" the no pooling estimates back towards the complete pooling mean, to an extent determined by the information in the data.

Next we formalize the model and consider some of its features and implications.

## Random Intercept Model

This model is a special case of a *random intercept* model in which covariates are categorical.  Note some consequences of this model.


$y_{ij}=\mu+\alpha_j+\varepsilon_{ij}$, where $\varepsilon_{ij} \overset{iid}{\sim} N(0,\sigma^2)$ $\perp$ $\alpha_j \overset{iid}{\sim} N(0,\tau^2)$

$E[y_{ij}]=E[\mu+\alpha_j+\varepsilon_{ij}]=\mu+0+0=\mu$
\begin{eqnarray*}
\text{Var}[y_{ij}]&=&E[(y_{ij}-E(y_{ij}))^2] \\
&=& E[(\mu+\alpha_j+\varepsilon_{ij}-\mu)^2] \\
&=& E[(\alpha_j+\varepsilon_{ij})^2] \\
&=& E[\alpha_j^2+2\alpha_j\varepsilon_{ij}+\varepsilon_{ij}^2] \\
&=& \tau^2+0+\sigma^2=\sigma^2+\tau^2
\end{eqnarray*}

##
For two observations in different groups j and j',
\begin{eqnarray*}
\text{Cov}(y_{ij},y_{i'j'})&=& E[(y_{ij}-E(y_{ij}))(y_{i'j'}-E(y_{i'j'}))] \\
&=& E(y_{ij}y_{i'j'})-\mu^2-\mu^2+\mu^2 \\
&=& E(y_{ij})E(y_{i'j'})-\mu^2=\mu^2-\mu^2=0
\end{eqnarray*}

\vspace{.1in}

For two observations in the same group j,
\begin{eqnarray*}
\text{Cov}(y_{ij},y_{i'j})&=& E[(y_{ij}-E(y_{ij}))(y_{i'j}-E(y_{i'j}))] \\
&=& E(y_{ij}y_{i'j})-\mu^2-\mu^2+\mu^2 \\
&=& E[(\mu+\alpha_j+\varepsilon_{ij})(\mu+\alpha_j + \varepsilon_{i'j})] \\
&=& E[\mu^2+\mu\alpha_j+\mu\varepsilon_{i'j}+\alpha_j\mu+\alpha_j^2+\alpha_j\varepsilon_{i'j}+ \\
& & ~~~~~\varepsilon_{ij}\mu+\varepsilon_{ij}\alpha_j+\varepsilon_{ij}\varepsilon_{i'j}] \\ 
&=& \mu^2 + 0 + 0 + 0 + \tau^2 + 0 + 0 + 0 + 0 -\mu^2=\tau^2
\end{eqnarray*}

## Intraclass Correlation

The correlation between two observations in the same group is then given by

\begin{eqnarray*}
\text{Corr}(y_{ij},y_{i'j})&=&\frac{\text{Cov}(y_{ij},y_{i'j})}{\sqrt(\text{Var}(y_{ij}))\sqrt(\text{Var}(y_{i'j}))} \\
&=& \frac{\tau^2}{\sigma^2+\tau^2}
\end{eqnarray*}

This motivates the use of random effects ANOVA to handle cases in which we expect subgroups of observations to be correlated (e.g., repeated measures or family studies).

##

It is often convenient to stack our observations into a long vector $Y_{N \times 1}$ organized by groups.  For simplicity in exposition, assume $n_j=n$ and the total sample size $N=nJ$. Then assuming $Y$ follows a multivariate normal distribution (which follows from our specification previously), we can express $$\text{Cov}(Y)=\sigma^2I_{N\times N} + \tau^2 \begin{pmatrix} J_n & 0 & \cdots & 0 \\ 0 & J_n & \cdots & 0 \\
\vdots & \vdots & \vdots & \vdots \\ 0 & 0 & \cdots & J_n \end{pmatrix}=I_J \otimes V,$$ where $V=\sigma^2I_n+\tau^2J_n$ and $J_n$ is an $n \times n$ matrix of 1's.

## Kronecker product

The *Kronecker product* is a convenient way to express patterned covariance matrices (among other things).  For matrices $A_{m \times n}$ and $B_{p \times q}$, the *Kronecker product* $A \otimes B=\begin{bmatrix}a_{11}B & \cdots & a_{1n}B \\ \vdots & \ddots & \vdots \\ a_{m1}B & \cdots & a_{mn}B \end{bmatrix}$.

Using a Kronecker product, we can succinctly express the block diagonal covariance matrix of all our observations when we have equal numbers of observations in each group.


## Estimation Methods

We briefly consider the following estimation methods for random intercept models.

  - Maximum likelihood (ML)
  
  - Restricted maximum likelihood (REML)
  
  - Empirical Bayes estimation
  

## Maximum Likelihood Estimation

We can also think of this formulation in the framework of the general linear mixed effects model, where $$y=X\beta+Zb+\varepsilon.$$ In the random effects ANOVA case, 

  - $X$ is just a column of 1's specifying the intercept $\beta=\mu$ 
  - $Z$ is a matrix of indicator variables indicating group membership 
  - Assume the random effects $b \sim N(0,G)$ where $G=\tau^2I$ and the errors $\varepsilon \sim N(0,R)$ where $R=\sigma^2I$
  
  The covariance matrix $\Sigma$ is then given by 
\begin{eqnarray*}
\Sigma&=&\text{Var}(y)=\text{Var}(X\beta+Zb+\varepsilon) \\
&=& \text{Var}(X\beta)+\text{Var}(Zb)+\text{Var}(\varepsilon)  \\
&=& Z\text{Var}(b)Z' + \text{Var}(\varepsilon) = ZGZ'+R =\tau^2ZZ'+\sigma^2I
\end{eqnarray*}

##

Assuming our $N$ outcomes follow the multivariate Gaussian distribution, our likelihood is given by $$\frac{1}{(2\pi)^\frac{N}{2}|\Sigma|^\frac{1}{2}}\exp\left(-\frac{1}{2}(y-X\beta)'\Sigma^{-1}(y-X\beta)\right),$$

and we often work with the log-likelihood, given by

\begin{eqnarray*}
\ell(y,\beta,\Sigma)&=&-\frac{1}{2}\left\{N\log(2\pi) + \log |\Sigma| + (y-X\beta)'\Sigma^{-1}(y-X\beta)   \right\} \\
&\propto& \log |\Sigma| + (y-X\beta)'\Sigma^{-1}(y-X\beta),
\end{eqnarray*}

which we then minimize (as I took the negative) in order to find the MLE.

## MLE for Simplest Case

Consider a one-sample setting in which we wish to estimate the sample mean $\mu$ and variance $\sigma^2$ using the model $y_{i}=\mu+\varepsilon_{i}$, $i=1,\ldots,n$, $\varepsilon_i \sim N\left(0,\sigma^2\right)$.

Then our log-likelihood is proportional to $n\log\sigma^2 + \frac{\sum(y_i-\mu)^2}{\sigma^2}$, and to find the MLE's of $\mu$ and $\sigma^2$, when you take derivatives and solve for zero, you obtain $\widehat{\mu}=\overline{y}$ and $\widehat{\sigma}^2=\frac{\sum (y_i-\overline{y})^2}{n}$. 

Typically we don't use the MLE to estimate $\sigma^2$ because of its well-known small-sample bias, instead using the unbiased estimate $s^2=\frac{\sum (y_i-\overline{y})^2}{n-1}=\frac{n}{n-1}\widehat{\sigma}^2$.

## MLE for Bike Data
 Recall our model
 
$y_{ij}=\mu+\alpha_j+\varepsilon_{ij}$, where $\varepsilon_{ij} \overset{iid}{\sim} N(0,\sigma^2)$ $\perp$ $\alpha_j \overset{iid}{\sim} N(0,\tau^2)$, where $y_{ij}$ indicates the passing distance between the car and the bike, and $\alpha_j$ represent effects of different distances between the bike and the curb. 

##

Instead of directly maximizing the log-likelihood in R, we can use the lme4 library to do the work for us.

```{r findmle, eval=FALSE}
load("~/Documents/GitHub/STA-410-610/PsychBikeData.RData")
library(lme4)
fit.ml=lmer(`passing distance` ~ (1 | kerb), REML=FALSE, data = PsychBikeData)
summary(fit.ml)
```

##

```{r findmle2, echo=FALSE}
load("~/Documents/GitHub/STA-410-610/PsychBikeData.RData")
library(lme4)
fit.ml=lmer(`passing distance` ~ (1 | kerb), REML=FALSE, data = PsychBikeData)
summary(fit.ml)
```


Our ML estimates of $(\mu,\tau^2,\sigma^2)$ for the bike data are $(\widehat{\mu},\widehat{\tau}^2,\widehat{\sigma}^2)=(1.540, 0.009, 0.137)$


## REML

REML (restricted or residual maximum likelihood) estimation is quite popular for variance component estimation. Features of REML estimation include the following

  - it is based on a likelihood function that only uses information that does not depend on fixed effects (we define new outcomes orthogonal to the mean)
  - it is generally less biased than ML estimates (and unbiased in certain special cases)
  
## REML for Simplest Case  

Recall our one-sample setting in which we wish to estimate the sample mean $\mu$ and variance $\sigma^2$ using the model $y_{i}=\mu+\varepsilon_{i}$, $i=1,\ldots,n$, $\varepsilon_i \sim N\left(0,\sigma^2\right)$.

REML estimates are often based on a full-rank set of error contrasts -- the basic idea is to retain the information in the data about the variance while eliminating the fixed effects. Because the residuals sum to zero in our simple model by definition, a full-rank set of residuals contains $n-1$ residuals for our mean-only model. If you've taken advanced linear models, you can show that the REML estimator maximizes the likelihood of a full rank set of error contrasts.

##

Because not everyone took STA 721, we'll consider a heuristic derivation. The full residuals $\varepsilon_i=y_i-\mu$ contain all the information in the likelihood about the variance parameter $\sigma^2$. 

Because the residuals are independent of the fixed effect $\mu$ and the sample mean $\overline{y}$ is independent of the sample variance, we can re-express our log likelihood to isolate the residual likelihood:

$$\ell(y,\mu,\sigma^2)=\ell(\varepsilon,\mu,\sigma^2)+\ell(\overline{y},\mu,\sigma^2)$$

We know $\widehat{\mu}=\overline{y}\sim N\left(\mu,\frac{\sigma^2}{n}\right)$ and so $\ell(\overline{y},\mu,\sigma^2)\propto \log \frac{\sigma^2}{n} + \frac{(\overline{y}-\mu)^2}{\frac{\sigma^2}{n}}$ which reduces to $\log \sigma^2 - \log n$ once we plug in the MLE $\overline{y}$ for $\mu$.

##

Then $$\ell(\varepsilon,\mu,\sigma^2) \propto n\log \sigma^2 + \frac{\sum (y_i-\mu)^2}{\sigma^2}-\log \sigma^2+\log n$$ which is proportional to $$(n-1)\log \sigma^2 + \frac{\sum (y_i-\mu)^2}{\sigma^2},$$ which looks just like our ML likelihood with the exception of the multiplier $n-1$ instead of $n$, and it's straightforward to show the maximum is $\widehat{\sigma}^2_{REML}=\frac{\sum (y_i-\mu)^2}{n-1}$.


  Because they are generally less biased than ML estimates, REML estimates are typically the default frequentist estimates provided by many software packages.


## REML Estimates for the Bike Data



```{r findreml}

fit.reml=lmer(`passing distance` ~ (1 | kerb), REML=TRUE, data = PsychBikeData)
summary(fit.reml)
```

Our REML estimates for the bike data are $(\widehat{\mu},\widehat{\tau}^2,\widehat{\sigma}^2)=(1.540, 0.012, 0.137)$. 

##

Note that the "REML" estimate of the mean $\mu$ is not obtained using the REML likelihood, which does not involve parameters in the linear predictor. We will return to this point later in the course when we deal with more complex models than this simple case, in which the mean estimate is $\overline{y}$ regardless of whether we use ML or REML for $\sigma^2$. In more complicated models, the estimates of mean parameters will be a function of variance parameters, and iterative methods are required.

## Empirical Bayes

When we have random effects in a model, the standard frequentist effects of these random quantities are called *empirical Bayes* estimates, regardless of whether we obtain other estimates using ML or REML.

##

Recall our group means formulation:

\begin{eqnarray*}
y_{ij}&=&\mu_j+\varepsilon_{ij}\\
\mu_1,\cdots,\mu_J &\overset{iid}{\sim}& N(\mu, \tau^2) \\
\varepsilon_{ij} &\overset{iid}{\sim} & N(0,\sigma^2).
\end{eqnarray*}

Suppose $(\mu, \tau^2, \sigma^2)$ are known exactly and consider estimating $\mu_j$ with an estimator that is a linear function of the group sample mean $\widehat{\mu}_j=a\overline{y}_j+b$. Then one can show that the MSE $E[(\mu_j-\widehat{\mu}_j)^2]$ is minimized if $a=\frac{\frac{n_j}{\sigma^2}}{\frac{n_j}{\sigma^2}+\frac{1}{\tau^2}}$ and $b=(1-a)\mu$, so that $\widehat{\mu}_j=w_j \overline{y}_j+(1-w_j)\mu$, where $w_j=\frac{\frac{n_j}{\sigma^2}}{\frac{n_j}{\sigma^2}+\frac{1}{\tau^2}}$
  
##

If we knew $(\mu, \tau^2,\sigma^2)$ this estimate would be the *Bayes estimate*; however, we do not know these parameters and are instead estimating them from the data, so that 

$\widehat{\mu}_j=\widehat{w}_j \overline{y}_j+(1-\widehat{w}_j)\widehat{\mu}$, where $\widehat{w}_j=\frac{\frac{n_j}{\widehat{\sigma}^2}}{\frac{n_j}{\widehat{\sigma}^2}+\frac{1}{\widehat{\tau}^2}}$

is called an *empirical Bayes estimate* because our unknown parameters have been replaced by "empirical" estimates from the data.
  
While this estimate is widely-used, it has several unsatisfactory qualities, including a standard variance estimate known to be an underestimate. This is great motivation for consideration of Bayesian approaches when formal comparisons among groups modeled with random effects are desired.

## EB Estimates of Group Means for Bike Data

```{r ebest}
table(PsychBikeData$kerb); mean(PsychBikeData$`passing distance`)
tapply(PsychBikeData$`passing distance`,PsychBikeData$kerb,mean)
coef(fit.ml)
```

##

```{r ebest2}
tapply(PsychBikeData$`passing distance`,PsychBikeData$kerb,mean)
coef(fit.reml)

```

Here we see only a slight shrinkage back towards the overall mean, due in large part to the large sample sizes within curb distances.




## Radon Redux: Adding a Grouping Factor

One important predictor of radon levels is the floor on which the measurement is taken:  basement ($x_i=0$) or first floor ($x_i=1$). Radon comes from underground and can enter more easily when the house is built into the ground; in addition, basements tend to have higher levels than ground floors.

First, we examine the complete-pooling regression, $y_{ij}=\alpha+\beta x_{ij} + \varepsilon_{ij}$, and the no-pooling regression $y_{ij}=\alpha_{j}+\beta x_{ij} + \varepsilon_{ij}$, where $\alpha_{j}$ is the mean log radon level from basement
measures of homes (indexed by i) in county $j$. 

The following plot shows the dashed lines $\widehat{y}=\widehat{\alpha}+\widehat{\beta} x$ for eight selected counties from the complete pooling model, and the solid lines
$\widehat{y}=\widehat{\alpha}_j+\widehat{\beta}x$ from no pooling model.

## No pooling and pooling

![Complete-pooling (dashed) and no-pooling (solid) regression fits to radon
data (Gelman and Hill)](figures/gelman2.jpg){width=100%}
  
## Interpretation

The estimates of $\beta$ (the association between floor of home and radon level) differ slightly for the two regressions, with $\widehat{\beta}=-0.61$ for the pooling model, and $\widehat{\beta}=-0.72$ for the no-pooling model. As we might expect, we tend to have higher radon levels in the basement (p<0.0001).  

Neither analysis is perfect.  The complete-pooling analysis ignores variation in radon levels between counties, which is undesirable because our goal is to identify counties with high-radon homes -- we can't pool away the main
research question!  The no-pooling analysis is also problematic -- for example the Lac Qui Parle County line is estimated based on just two data points.

## Multilevel model

We will start with a simple multilevel model, $y_{ij}=\gamma_0 +
\alpha_{j}+\beta x_{ij} + \varepsilon_{ij}$, where now $\alpha_{j} \sim N(0,\tau^2)$ and $\varepsilon_{ij} \sim N(0, \sigma^2)$.  We fit this model using the lmer() function in the lme4 package.

This model can also be parameterized $y_{ij} \sim N(\alpha_{j}+\beta x_{ij}, \sigma^2),$ 

$\alpha_j \sim N\left(\gamma_0,\tau^2 \right)$

## Code to fit models

Data: [srrs2.dat](data/srrs2.dat)

```{r readsetup}
srrs2 <- read.table ("data/srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]   # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

```

```{r fitsimplemodels, eval=FALSE}
###pooled model
lm.pooled<-lm(formula=y~x)
summary(lm.pooled)
###unpooled  model
lm.unpooled<-lm(formula=y~x+factor(county)-1)
summary(lm.unpooled)
```

##

```{r fitsimplemodels2, echo=FALSE}
###pooled model
lm.pooled<-lm(formula=y~x)
summary(lm.pooled)
```

##

```{r fitsimplemodels3, echo=FALSE}
###unpooled  model
lm.unpooled<-lm(formula=y~x+factor(county)-1)
summary(lm.unpooled)
```


##

```{r lmer, eval=FALSE}
#basic MLM with just random intercept for county
library(lme4)
M1<-lmer(y~x+(1|county),REML=FALSE)
summary(M1)
coef(M1)
```


##

```{r lmer2, echo=FALSE}
library(lme4)
#basic MLM with just random intercept for county
M1<-lmer(y~x+(1|county),REML=FALSE)
summary(M1)
coef(M1)
```


```{r M2}
library(lme4)
#basic MLM with just random intercept for county, fit using REML
M2<-lmer(y~x+(1|county))
```

## Output from MLM
```{r M2out}
summary(M2)
```

<small> The ratio of county-level variance to total variance is an estimate of the correlation of within-county measures. Here $\hat{\rho}=\frac{.1077}{.1077+.5709}=0.16$. </small>

## County-level estimates
```{r M2colevel}
coef(M2)
```

The first column gives estimates of $\alpha_j$, and the second column gives the estimate of $\beta$ (does not vary over county $j$ according to the model we specified).

## Your Job!

Examine estimates from complete pooling, no pooling, and the multilevel model for counties in Pennsylvania (PA), with a focus on how the shrinkage varies according to sample size.



